---
title: "Winter Crops Optimisation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Winter Crops Optimisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ram)
```

## Problem Statement

A Mediterranean farm plans to allocate land to 4 winter crops:
- Gramineae: Oat, Barley
- Legumes: Fava bean, Lupin

## Scenario:

- Total land available: 15 ha
- Total labor hours: 350 hours
- Total nitrogen: 500 kg
- Oat: 2 ≤ area ≤ 8 ha
- Barley: 1 ≤ area ≤ 6 ha
- Fava bean: 3 ≤ area ≤ 7 ha
- Lupin: 2 ≤ area ≤ 5 ha

### Objective: Maximize total gross margin (profit)

| Crop      | N fertilizer need (kg $N/ha$) | Work labor ($h/ha$) | Gross margin ($€/ha$) |
| :-------- | :-------------------------: | :----------------: | :-----------------: |
| Oat       |              80             |         20         |         400         |
| Barley    |             100             |         25         |         450         |
| Lupin     |              0              |         15         |         350         |
| Fava bean |              30             |         30         |         500         |


- **Objective**: Maximize total gross margin.

## Step 1: Define Resource Constraints (flexible min\max for each crop)

```{r}
# Step 1: Define Resource Constraints (flexible min/max for each crop)
resources <- data.frame(
  resource     = c("land", "labor","nitrogen"),
  availability = c(15, 350, 500),
  direction    = c("<=", "<=","<=")
)
print(resources)
```

### Step 2: Define Activities (Crop Choices)

```{r}
# Step 2: Define Activities 
activities <- data.frame(
  activity  = c("oat", "barley",  "lupin", "fava"),
  land      = c(1, 1, 1, 1),
  labor     = c(20, 25, 15, 30),
  nitrogen  = c(80, 100, 0, 30),
  objective = c(400, 450, 350, 500)
)
print(activities)
```

### Step 3: Build and Solve the Model

```{r}
# Step 3: Build and Solve the Model
model <- create_ram_model(resources, activities)
solution <- solve_ram(model, direction = "max")
```

### Step 4: Results

```{r}
summary_ram(solution)
cat("Maximum gross margin (EUR):", solution$objective_value, "\n")
```


```{r, fig.width=10, fig.height=5}
plot_ram(solution)
```


