---
title: "Winter Crops Optimization with Area Bounds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Winter Crops Optimization with Area Bounds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(ram)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Problem Statement

A Mediterranean farm has 15 ha of land, 350 h of labor and 500 kg of available nitrogen.
It can grow four winter crops:

Oat (a gramineae)

Barley (a gramineae)

Fava bean (a legume)

Lupin (a legume)

Crop data (per ha):

| Crop      | N fertilizer (kg) | Labor (h) | Gross margin (€) |
| :-------- | :---------------: | :-------: | ---------------: |
| Oat       |         80        |     20    |              400 |
| Barley    |        100        |     25    |              450 |
| Lupin     |         0         |     15    |              350 |
| Fava bean |         30        |     30    |              500 |


We wish to maximize total gross margin subject to resource constraints.
Furthermore, each crop has a required area range (ha):

Oat: 2 ≤ area ≤ 8

Barley: 1 ≤ area ≤ 6

Fava: 3 ≤ area ≤ 7

Lupin: 2 ≤ area ≤ 5

## Step 1: Define Resource Constraints (flexible min/max for each crop)

```{r}
# Step 1: Define Resource Constraints (flexible min/max for each crop)
resources <- define_resources(
  resources = c(
    # global totals
    "land", "labor", "nitrogen",
    # per‐crop bounds
    "oat_min",   "oat_max",
    "barley_min","barley_max",
    "lupin_min", "lupin_max",
    "fava_min",  "fava_max"
  ),
  availability = c(
    # totals
     15,    350,    500,
    #  bounds
      2,      8,
      1,      6,
      3,      7,
      2,      5
  ),
  direction = c(
    # totals all <=
    "<=",   "<=",   "<=",
    # minimums >=
    ">=",   "<=",
    ">=",   "<=",
    ">=",   "<=",
    ">=",   "<="
  )
)

print(resources)
```

### Step 2: Define Activities (Crop Choices)

```{r}
# Step 2: Define Activities
activity_requirements_matrix <- matrix(
  c(
    # land usage (ha/ha)
     1, 1, 1, 1,
    # labor   (h/ha)
    20,25,15,30,
    # nitrogen (kg N/ha)
    80,100, 0,30,
    # bounds rows for technical matrix:
    #   for oat_min  (1 for oat else 0)
     1,0,0,0,
    #   for oat_max
     1,0,0,0,
    # barley_min
     0,1,0,0,
    # barley_max
     0,1,0,0,
    # fava_min
     0,0,1,0,
    # fava_max
     0,0,1,0,
    # lupin_min
     0,0,0,1,
    # lupin_max
     0,0,0,1
  ),
  nrow = nrow(resources),
  byrow = TRUE,
  dimnames = list(
    resources$resource,
    c("oat", "barley", "lupin", "fava")
  )
)

objective <- c(oat = 400, barley = 450, lupin = 350, fava = 500)

activities <- define_activities(
  activities = colnames(activity_requirements_matrix),
  activity_requirements_matrix = activity_requirements_matrix,
  objective = objective
)

activities
```

### Step 3: Build and Solve the Model

```{r}
# Step 3: Build and Solve the Model
model <- create_ram_model(resources, activities)
solution <- solve_ram(model, direction = "max")
```

### Step 4: Results

```{r}
dat <- data.frame(
  activity = names(solution$optimal_activities),
  level    = round(as.numeric(solution$optimal_activities), 1)
)

DT::datatable(
  dat,
  rownames = FALSE,
  options = list(
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '2cm', targets = 0),  # first column (activity)
      list(width = '2cm', targets = 1)   # second column (level)
    )
  )
) |>
  DT::formatRound(
    columns = "level",
    digits  = 2
  )
```

### Step 4: Results

```{r}
summary_ram(solution)
cat("Maximum gross margin (EUR):", solution$objective_value, "\n")
```

```{r, fig.width=10, fig.height=5}
plot_ram(solution)
```

## Scenario Analysis

Below we explore what happens if we vary total nitrogen and labour availability.


```{r}
library(ram)

# 1) Re‐load your data.frames from the vignette/example:
resources   # data.frame with columns resource, availability, direction
activities  # data.frame with columns activity, land, labor, nitrogen, …, objective

# 2) Build the ram specs:
res0 <- define_resources(
  resources    = resources$resource,
  availability = resources$availability,
  direction    = resources$direction
)

# NOTE THE TRANSPOSE HERE!
tech_mat <- t( as.matrix(
   activities[, setdiff(names(activities), c("activity","objective"))]
) )

act0 <- define_activities(
  activities                   = activities$activity,
  activity_requirements_matrix = tech_mat,
  objective                    = activities$objective
)

# 3) Define your scenario grid:
scenarios <- list(
  nitrogen = c(300, 400, 500),
  labor    = c(300, 350)
)

# 4) Run them:
scen_df <- run_scenarios(res0, act0, scenarios, direction = "max") 
print(scen_df)
```


```{r fig.alt="Line chart showing how maximum gross margin varies with nitrogen levels under different labor availabilities."}
library(ggplot2)
ggplot(scen_df, aes(x = factor(nitrogen), y = objective, color = factor(labor), group = factor(labor))) +
  geom_line() + geom_point() +
  labs(x = "Nitrogen (kg)", y = "Max Gross Margin (€)", color = "Labor (h)")
```
